---
title: "dmeav02 Usage"
author: "Francisco Richter M."
date: "June 1, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dmea02)
```


## tree simulation and consistency 

functions

1. ```sim.tree``` to simulate trees
2. ```phylo2vectors``` to convert a phylo into the set of vectors characterizing the tree

```{r}
s = sim.tree(ct=6)
plot(s$phylo)
s$tree
phylo2vectors(s$phylo)
all.equal(s$tree$wt,phylo2vectors(s$phylo)$wt)
all.equal(s$tree$E,phylo2vectors(s$phylo)$E)
all.equal(s$tree$n,phylo2vectors(s$phylo)$n)
```


to do: 

1. ```phylo2vectors```does not give the vector S corresponding to topology. add it
2. remove the root on vector outputs, keep consistency 
3. arenge the vectors2phylo function and make a lot simpler the ```sim.tree``` using it
4. add the log in an log output file

### Tree manipulation

```{r}
s = sim.tree(ct=2,seed=5)
plot(s$phylo)
s2=phylo2vectors(s$phylo)
s2
s$tree
up = update.tree(s2,0.5,1.5)
up
```

1. Doing update tree I realized that I should be more carefully in analysis when we lost crown time.
2. phylo2vectors really needs to take into account topology


#### more about consistency

```{r}
s <- sim.tree()
plot(s$phylo)
plot(vectors2phylo(s$tree))
all.equal(s$phylo,vectors2phylo(s$tree))
```

## The reconstruction algorithm

```{r}
s = sim.tree(ct=6)
plot(s$phylo.extant)
rec = rec.tree(tree=s$tree.extant,pars=c(0.8,0.1,40))
plot(vectors2phylo(rec))
s$tree.extant
rec
```

### let´s observe one estimation
```{r, sim1}
s = sim.tree(seed = 1)
mle.tree(s$tree)
tree = s$tree.extant
st = sim.srt(tree,pars=c(0.8,0.1,40),n_trees=100)
mle.st(st)

```


### How good is the last iteration of the MCEM algorithm?

```{r, eval=FALSE}
library(dmea)
n_sim = 537
n_trees = 10
MP = matrix(nrow=n_sim,ncol=3)
RP = matrix(nrow=n_sim,ncol=3)
p = proc.time()
for(i in 1:n_sim){
  est = sim.est(n_trees=n_trees,pars=c(0.8,0.1,40),seed=i)
  RP[i,] = est$real
  MP[i,] = est$est
}
print(proc.time()-p)
par_est_vis(P=MP,par=1,PR=RP)
par_est_vis(P=MP,par=2,PR=RP)
par_est_vis(P=MP,par=3,PR=RP)
```
