---
title: "lightnings segments experiment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(ggplot2)
```


## First we create the functions we are going to use

```{r}
#simulate segments, the output are waiting times, topology and observed waiting times
sim.seg <- function(lambda,p,n=1){
  to = NULL
  t = NULL
  obs = NULL
  for(i in 1:n){
    to1=0
    while(to1 == 0){
      to1 = rbinom(1,1,p)
      to = c(to,to1)
      t = c(t,rexp(1,lambda))
    }
    obs = c(obs,sum(t)-sum(obs))
  }
  return(list(t=t,to=to,obs=obs))
}

# simulate missing data given the observed segment
rec.seg <- function(obs,lambda,p){
  to=NULL
  t=NULL
  for(i in 1:length(obs)){
    Dt = obs[i]
    n = rpois(1,(1-p)*lambda*Dt)
    it = sort(runif(n,min=0,max=Dt)) #inter-arrival times
    t = c(t,diff(c(0,it,Dt)))
    to = c(to,rep(0,times=n),1)
  }
  return(list(t=t,to=to))
}

#calculate the mle for a set of segments
mle.ss <- function(S){
  m = length(S)
  n = vector(mode='numeric',length = m)
  d = vector(mode='numeric',length = m)
  dt = vector(mode='numeric',length = m)
  for(i in 1:m){
    rec = S[[i]]
    d[i] = length(rec$t)
    n[i] = sum(rec$to)
    dt[i] = sum(rec$t)
  }
  lambda.mle = sum(d)/sum(dt)
  p.mle = sum(n)/sum(d)
  return(list(lambda.mle=lambda.mle,p.mle=p.mle))
}

# monte-carlo sampling of size m
sim.srs <- function(Dt,lambda,p,m){ 
  Rec = vector(mode = 'list',length = m)
  d = vector(mode='numeric',length = m)
  dt = vector(mode='numeric',length = m)
  for(j in 1:m){
    rec=rec.seg(Dt,lambda,p)
    Rec[[j]] = rec
  }
  return(Rec)
}



```


## After that we can do some examples
### Exercises 1. a random segment

```{r}
lambda = 0.8
p = 0.5
#create random segment
seg = sim.seg(lambda = 0.8,p = 0.5)
#segment
seg 
#observed
Dt = seg$obs
m=10000
# monte-carlo sampling
R = sim.srs(Dt,lambda,p,m)
# mle
mle.ss(R)

# now with more segments
seg = sim.seg(lambda = 0.8,p = 0.5,n=25)
Dt=seg$obs
srs = sim.srs(Dt,lambda,p,m)
mle.ss(srs)

```



## Exercises 2. How good is the last iteration of MCEM
### To perform this expetiment we run the following rutine l times
1. simulate tree and get observed data
2. MC sampling for the missing part
3. estimate MLE and store

```{r, ex0001s}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=l)
P=vector(mode = 'numeric',length=l)
lambda=0.8
p=0.8
m=100
for(k in 1:l){
  seg = sim.seg(lambda = lambda, p = p,n=40)
  Dt = seg$obs
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = mle$lambda.mle
  P[k] = mle$p.mle
}
print(proc.time()-time)
summary(L)
summary(P)

```

Quite accurate estimations, now let´s observe another with a lot of missing data

```{r, ex001s}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=l)
P=vector(mode = 'numeric',length=l)
lambda=0.8
p=0.01
m=100
for(k in 1:l){
  seg = sim.seg(lambda = lambda, p = p,n=40)
  Dt = seg$obs
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = mle$lambda.mle
  P[k] = mle$p.mle
}
print(proc.time()-time)
summary(L)
summary(P)
```


In general they are not biased and accurate. We can change the mc sampling size $m$, the length of the tree $n$ or the parameters, but they are more or less working fine. Now let´s see how the MCEM works. 


# MCEM 



```{r}
mcem.light <- function(Dt,lambda,p,nit=100){
  m=2
  L=vector(mode = 'numeric',length=l)
  P=vector(mode = 'numeric',length=l)
  for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
  m=m+1
  }
  return(list(L=L,P=P))
}
```


From now onwards I perform several MCEM routines to observe how is working with different input parameters. 


```{r, mcem1}
time = proc.time()
l=300
lambda=0.8
p=0.5
seg = sim.seg(lambda = lambda, p = p,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem2}
time = proc.time()
l=300
lambda=0.8
p=0.5
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem3}
time = proc.time()
l=300
lambda=0.8
p=0.5
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem4}
time = proc.time()
l=300
lambda=0.8
p=0.5
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem5}
time = proc.time()
l=300
lambda=0.8
p=0.5
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem6}
time = proc.time()
l=300
lambda=0.8
p=0.1
seg = sim.seg(lambda = lambda, p = p,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem7}
time = proc.time()
l=300
lambda=0.8
p=0.1
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem8}
time = proc.time()
l=300
lambda=0.8
p=0.9
seg = sim.seg(lambda = lambda, p = p,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```


```{r, mcem9}
time = proc.time()
l=300
lambda=0.8
p=0.5
#seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```

```{r, mcem10}
time = proc.time()
l=300
lambda=0.8
p=0.99
seg = sim.seg(lambda = 0.8, p = 0.5,n=50)
Dt = seg$obs
mcem = mcem.light(Dt,lambda,p,m)
print(proc.time()-time)

qplot(1:l,mcem$L) + geom_hline(yintercept = lambda)
qplot(1:l,mcem$P) +  geom_hline(yintercept = p)


```

```{r, extra}
m=100
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.5)
P*L
length(Dt)/sum(Dt)
```

```{r, extra2}
m=10
lambda=1.2
p=0.1
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.5)
P*L
length(Dt)/sum(Dt)
```

```{r}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=l)
P=vector(mode = 'numeric',length=l)
lambda=0.8
p=0.5
m=10
seg = sim.seg(lambda = 0.8, p = 0.5,n=100)
Dt = seg$obs
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.5)
qplot(1:length(P),P*L) +  geom_hline(yintercept = 0.5*0.8)


```

```{r}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=l)
P=vector(mode = 'numeric',length=l)
lambda=0.8
p=0.5
m=10
seg = sim.seg(lambda = 0.8, p = 0.5,n=100)
Dt = seg$obs
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.5)
qplot(1:length(P),P*L) +  geom_hline(yintercept = 0.5*0.8)


```



```{r}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=l)
P=vector(mode = 'numeric',length=l)
lambda=0.8
p=0.5
m=100
seg = sim.seg(lambda = 0.8, p = 0.5,n=100)
Dt = seg$obs
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k] = lambda = mle$lambda.mle
  P[k] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.5)
qplot(1:length(P),P*L) +  geom_hline(yintercept = 0.5*0.8)


```



```{r}
time = proc.time()
l=1000
L=vector(mode = 'numeric',length=(l+1))
P=vector(mode = 'numeric',length=(l+1))
L[1] = lambda=0.8
P[1] = p = 0.5
m=10
seg = sim.seg(lambda = 0.8, p = 0.1,n=100)
Dt = seg$obs
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k+1] = lambda = mle$lambda.mle
  P[k+1] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.1)
qplot(1:length(P),P*L) +  geom_hline(yintercept = 0.1*0.8)
qplot(1:(length(P)-1),diff(L)) 

```


```{r}
time = proc.time()
l=100
L=vector(mode = 'numeric',length=(l+1))
P=vector(mode = 'numeric',length=(l+1))
L[1] = lambda=0.8
P[1] = p = 0.5
m=100
seg = sim.seg(lambda = 0.8, p = 0.1,n=100)
Dt = seg$obs
for(k in 1:l){
  srs = sim.srs(Dt,lambda,p,m)
  mle=mle.ss(srs)
  L[k+1] = lambda = mle$lambda.mle
  P[k+1] = p = mle$p.mle
}
print(proc.time()-time)

qplot(1:length(P),L) + geom_hline(yintercept = 0.8)
qplot(1:length(P),P) +  geom_hline(yintercept = 0.1)
qplot(1:length(P),P*L) +  geom_hline(yintercept = 0.1*0.8)
qplot(1:(length(P)-1),diff(L)) 

```


### NHPP 


```{r}
# NHPP
nhpp.missing <- function(pars,ct=15){
  lambda_0 = pars[1]
  mu_0 = pars[2]
  t_max     <- ct
  t         <- 0
  s         <- 0
  lambda <- function(t) lambda_0*(1-exp(-mu_0*(ct-t)))
  Lambda <- function(tupper) integrate(f = lambda, lower = 0, upper = tupper)$value
  Lambda_inv <- function(s){
    v <- seq(0,t_max+3, length.out = 1000)
    min(v[Vectorize(Lambda)(v)>=s])
  }
  X <- numeric(0)
  while(t <= t_max){
    u <- rexp(1)
    s <- s + u
    if(s > Lambda(ct)) break
    t <- Lambda_inv(s)
    X <- c( X, t)
  }
return(X)
}
```



